### 2016-03-29 msmfs.tex
First commit

%rubber: module pdflatex

\documentclass[11pt,a4paper]{article}

\usepackage[sans]{tcc_doc} % SDP style file
\usepackage[english]{babel}
\usepackage{amssymb,amsmath} % for the math symbols
\usepackage{geometry}
\usepackage{marginnote}
\usepackage{datetime}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{mathtools}
\graphicspath{ {images/} }

% TODO:
% 1. Check if all new symbols have been explained when used for the first
% time.
% 2. Check if all symbols are actually being used.
% 3. Check the symbol definitions: rates, numbers, etc. to use the same main letter
% 4. Is there a way to use proper referencing of applicable and reference
% documents?

%%%%%%%%% START OF USER SETTINGS %%%%%%%%%%%%%%%%%%

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Implementation of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-SDP-151123-2\xspace}
% Context
\newcommand{\context}{(SDP.PIP)}
% Revision
\newcommand{\revision}{2\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwel\\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Draft\xspace}

% The definitions below are used for the signature tables in the preamble of
% the document
\newcommand{\firstsigname}{Name 1 - to be filled in}
\newcommand{\firstsigdesignation}{Designation 1 - to be filled in}
\newcommand{\firstsigaffiliation}{Affiliation 1 - to be filled in}

\newcommand{\secondsigname}{Name 2 - to be filled in}
\newcommand{\secondsigdesignation}{Designation 2 - to be filled in}
\newcommand{\secondsigaffiliation}{Affiliation 2 - to be filled in}

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
  \end{tabularx}
}

%%%%%%%%%%%%% END OF USER SETTINGS %%%%%%%%%%%%%%%%%%%

% Table with signatures
\newcommand{\signaturetable}[3]{
  \begin{tabularx}{\textwidth}{|X|X|X|}
    \hline
    Name & Designation & Affilitation\\
    \hline
    {#1} & {#2} & {#3}\\
    \hline
    Signature \& Date: & & \\
    & & \\
    & & \\
    \hline
  \end{tabularx}
}

% Table with affiliations
\newcommand{\organisationtable}{
  \begin{center}
    \sffamily{\bf ORGANISATION DETAILS}\end{center}
  \begin{table}[htb]
    \centering
    \begin{tabular}[htb]{|l|l|}
      \hline
      Name & Science Data Processor Consortium\\
      \hline
    \end{tabular}
  \end{table}
}

%%%%%%%%%%%%%%%%%%%% START YOUR DOCUMENT HERE %%%%%%%%%%%%%%%%%

\newcommand{\subsubsubsection}[1]{\noindent{\bf{#1}}}

%%%% Please DO USE the symbol definitions listed below to preserve
%%%% consistency.

%%%% If you need a new symbol, please add it here. If really want to modify
%%%% one, do so, just take care to do a search-and-replace for the entire document
% Symbol definitions
\newcommand{\bytes}{b} % Length of a complex number in bytes
\newcommand{\cbytes}{b} % Length of a real number in bytes
\newcommand{\rbytes}{b} % Length of a complex number in bytes
\newcommand{\bl}{B} % Baseline length
\newcommand{\maxbl}{B_{\mathrm{max}}} % Maximum baseline length
\newcommand{\membw}{B_\mathrm{mem}} % Memory bandwidth
\newcommand{\pciebw}{B_\mathrm{PCIe}} % PCI-e bandwidth
\newcommand{\viscompr}{C} % Compression factor of visibilities for gridding
                          % for facets
\newcommand{\vism}{\mathbb{D}} % Matrix of observed visibilities
\newcommand{\da}{D_{\mathrm A}} % Extent of aperture
\newcommand{\ds}{D_{\mathrm s}} % Diameter of antennas/stations
\newcommand{\reff}{f_\mathrm{ref}} % Reference frequency
\newcommand{\fcc}{F_{\mathrm{cc}}} % Convolution kernel computational cost
\newcommand{\fci}{F_{\mathrm{ci}}} % Factor between bandwidth and FLOPS (bytes/ops)
\newcommand{\fgrid}{F_{\mathrm{grid}}} % Gridding cost
\newcommand{\fpatch}{f_\mathrm{patch}} % Fraction of the maximum baseline below which the uv coverage is almost filled
\newcommand{\fpr}{F_\mathrm{pr}} % Cost for phase rotation
\newcommand{\gainm}{\mathbb{G}} % 2x2 block diagonal matrix of complex gains
% To change to N_something
\newcommand{\modskyvism}{\mathbb{M}} % Matrix of model sky visibilities
\newcommand{\skymodbufsize}{M_\mathrm{buf,skymod}} % Size of the sky model buffer
\newcommand{\visbufsize}{M_\mathrm{buf,vis}} % Size of the visibility buffer
\newcommand{\visbufsizespec}{M^\mathrm{spec}_\mathrm{buf,vis}} % Size of the visibility buffer needed for spectral line processing
\newcommand{\visbufsizecont}{M^\mathrm{cont}_\mathrm{buf,vis}} % Size of the visibility buffer needed for continuum imaging
\newcommand{\bufsize}{M_\mathrm{cu,buf}} % Size of buffer
\newcommand{\slowmemsize}{M_\mathrm{cu,pool}} %	Size of slow working memory
\newcommand{\cumemsize}{M_\mathrm{cu,work}} % Size of working memory of compute unit
\newcommand{\imgsize}{M_\mathrm{image}} % Size of image (bytes)
\newcommand{\imgsizeoneax}{M_\mathrm{image,ax}}	% Image size along one axis
\newcommand{\nbyteperpix}{N_\mathrm{byte,pix}} % Number of bytes per pixel
%\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} % Number of bytes per pixel
\newcommand{\targgridsize}{M_\mathrm{target}} %	Target grid size (per facet)
\newcommand{\uvgridsize}{M_{uv,\mathrm{grid}}} % Size of the uv-grid (bytes)
\newcommand{\vissize}{M_\mathrm{vis}} % Size of visibility data (bytes)
\newcommand{\wgridsize}{M_\mathrm{weight\_grid}} % Size of weight grid (bytes)
\newcommand{\gridcachesize}{M_{w,\mathrm{cache}}} % Cache size of
                                % gridding kernel
\newcommand{\ntel}{N_\mathrm{a}} % Number of antennas or stations in the array
\newcommand{\nateam}{N_\mathrm{A}} % Number of A-team sources
\newcommand{\ngridpakern}{N_\mathrm{AA}} % Number of grid points for A-kernel support (along one axis)
\newcommand{\naccvis}{N_\mathrm{apv}} % Number of accesses per visibility
\newcommand{\navg}{N_\mathrm{avg}} % Number of averages
\newcommand{\nbins}{N_\mathrm{bin}} % Number of bins
\newcommand{\nbeam}{N_\mathrm{beam}} % Number of beams simultaneously observed by the array
\newcommand{\rma}{N_{rma}} % Number of flops in real multiply and add
\newcommand{\cma}{N_{cma}} % Number of flops in complex multiply and add
\newcommand{\nbits}{N_\mathrm{bit}} %	Number of bits
\newcommand{\nbl}{N_\mathrm{bl}} % Number of baselines
\newcommand{\nbyteacc}{N_\mathrm{bpa}} % Number of bytes per access
\newcommand{\nbyte}{N_\mathrm{byte}} % Number of bytes
\newcommand{\nch}{N_\mathrm{ch}XXXWANT TO USE NF HERE} % Number of channels
\newcommand{\ncompu}{N_\mathrm{cu}} % Number of compute units
\newcommand{\ndatait}{N_\mathrm{datait}} % Number of data items required per source (e.g. postion)
\newcommand{\ngridp}{N_\mathrm{cvff}} % Number of grid points along one axis for the gridding convolution function
\newcommand{\nfch}{N_\mathrm{f}} % Number of frequency channels
\newcommand{\nfchcorr}{N_\mathrm{f,corr}} % Number of frequency channels output by the correlator
\newcommand{\nfnosmear}{N_\mathrm{f,smear}} % Number of frequency channels set by smearing limit
\newcommand{\nfchgrid}{N_\mathrm{f,grid}} % Number of frequency channels to be gridded
\newcommand{\nfchdegrid}{N_\mathrm{f,de-grid}} % Number of frequency channels to be de-gridded
\newcommand{\nfchout}{N_\mathrm{f,out}} % Number of output frequency channels
                                % (to be FFT-ed and reprojected or output from
                                % the CSP pulsar timing backend
\newcommand{\nfchdist}{N_\mathrm{f,distribute}} % Minimum number of frequencies needed to maintain distributability
\newcommand{\nfkernel}{N_\mathrm{f,kernel}} %Number of convolution kernels needed to cover frequency axis for a particular baseline
\newcommand{\nfacet}{N_\mathrm{facet}} % Number of facets along one axis
\newcommand{\facetoverlap}{r_\mathrm{facet}} %Facet overlap overhead
\newcommand{\nflop}{N_\mathrm{FLOP}} % Number of floating point operations
\newcommand{\nflopsamp}{N_\mathrm{FLOPsamp}} % Number of floating point
                                % operations per sample
\newcommand{\nflopvis}{N_\mathrm{FLOPvis}} % Number of floating point
                                % operations per visibility 
\newcommand{\npixgrid}{N_\mathrm{grid}} % Number of pixels in the grid (i.e. the square of the linear dimension)
\newcommand{\ngridwkern}{N_\mathrm{GW}} % Number of grid points for w-kernel
                                % support (along one axis)
\newcommand{\nit}{N_\mathrm{it}} % Number of iterations
\newcommand{\nawkern}{N_\mathrm{kernel}} % Linear size of combined A and w-kernels
\newcommand{\nmaj}{N_\mathrm{major}} % Number of major CLEAN cycles
\newcommand{\nmin}{N_\mathrm{minor}} % Number of minor CLEAN cycles
\newcommand{\nselfcal}{N_\mathrm{SelfCal}} %Number of self cal loops
% TODO: Factors should have another letter than N
\newcommand{\facgridoffdiagmuller}{N_\mathrm{mm}} % Factor to account for
                                % gridding of off-diagonal Mueller terms
% TODO: Factors should have another letter than N
\newcommand{\nops}{N_\mathrm{ops}} % Number of operations
\newcommand{\npatchpix}{N_\mathrm{patch,pix}} % Number of pixels of PSF to be used in minor cycle decorrelation
\newcommand{\npix}{N_\mathrm{pix}} % Number of pixels along one axis of a grid
\newcommand{\npixfacet}{N_\mathrm{pix,facet}} % Number of pixels along one axis of a single image facet
\newcommand{\npol}{N_\mathrm{pol}} % Number of polarisation products
\newcommand{\nsubint}{N_\mathrm{subint}} % Number of sub-integrations
\newcommand{\ntslot}{N_t} % Number of time slots
\newcommand{\ntaylor}{N_\mathrm{Tt}} % Number of Taylor terms
\newcommand{\ntslotave}{N_{t,\mathrm{av}}} % Number of time slots that are averaged (in de-mixing)
\newcommand{\nvis}{N_\mathrm{vis}} % Number of visibilities
\newcommand{\nvisrate}{R_{vis}} % Number of visibilities per second\\
% TODO: what kind of qfac?
\newcommand{\qfacbw}{Q_\mathrm{bw}} % Quality factor
\newcommand{\qfacreuse}{Q_\mathrm{fcv}} % Factor to allow for the reuse of functions between neighbouring frequency channels
\newcommand{\qfacfov}{Q_\mathrm{FoV}} % Quality factor defining how much bigger the diameter of the field of view is than the first zero of the Airy function
\newcommand{\facgridovers}{Q_\mathrm{GCF}} % Gridding oversampling factor
\newcommand{\facmsmf}{Q_\mathrm{MSMF}} % Factor to account for multiple subtractions (multi-frequency multi-scale CLEAN)
\newcommand{\qfacbeamovers}{Q_\mathrm{pix}} % Quality factor defining the
                                % oversampling of the synthesised beam
% TODO: what kind of qfac?
\newcommand{\qfac}{Q_t} % Quality factor
\newcommand{\corrfacw}{Q_w} % Correction factor for w values (0 to 1)
\newcommand{\earthrad}{R_\mathrm{Earth}} % Radius of the Earth ($6400\times 10^3$\,m)
\newcommand{\comprate}{R} % Compute rate (FLOPS)
\newcommand{\compratespec}{R^\mathrm{spec}} % Compute rate for spectral line processing
\newcommand{\compratecont}{R^\mathrm{cont}} % Compute rate for continuum imaging
\newcommand{\compratefast}{R^\mathrm{fast}} % Compute rate for fast imaging
\newcommand{\covm}{\mathbb{R}} % Covariance matrix
\newcommand{\covmfilt}{\mathbb{R}_\mathrm{filt}} % Filtered covariance matrix
\newcommand{\bwtoworkmem}{R_\mathrm{bw,mem}} % Bandwidth to working memory
\newcommand{\maxmembw}{R_\mathrm{bw,max}} % Maximum memory bandwidth of
                                % compute unit
\newcommand{\convkernflops}{R_\mathrm{CCF}} % Convolution kernel compute rate (FLOPS)
\newcommand{\fftflops}{R_\mathrm{FFT}} % FFT compute rate (FLOPS)
\newcommand{\flagflops}{R_\mathrm{flag}} % FLOP rate to do flagging on input visibilities
\newcommand{\griddingflops}{R_\mathrm{grid}} % Gridding compute rate (FLOPS)
\newcommand{\phaserotflops}{R_\mathrm{PR}} % Phase rotation compute rate (FLOPS)
\newcommand{\reprojflops}{R_\mathrm{RP}} % Re-projection compute rate (FLOPS)
\newcommand{\rcorr}{R_\mathrm{s}} % Correlator output rate (samples/sec)
\newcommand{\peakflop}{R_\mathrm{peak}} % Peak FLOPS of compute unit
% TODO - verschil met max mem bw of compute unit?
% TODO - verschil memory bw and bw to buffer?
\newcommand{\IO}{{R_\mathrm{IO}}} %IO rate
\newcommand{\maxiobwbuff}{R_\mathrm{bw,I/O,max}} % Maximum I/O bandwidth of compute unit to buffer
\newcommand{\bwbuff}{R_\mathrm{bw,I/O}} % Bandwidth to buffer
\newcommand{\bwbuffcont}{R^\mathrm{cont}_\mathrm{bw,I/O}} % Bandwidth to
                                % buffer for cotinuum imaging
\newcommand{\bwbuffspec}{R^\mathrm{spec}_\mathrm{bw,I/O}} % Bandwidth to
                                % buffer for spectral line processing
\newcommand{\nsource}{N_\mathrm{source}} % Number of sources
\newcommand{\telsensit}{S} % Telescope sensitivity
\newcommand{\corrdumpt}{t_\mathrm{dump}} % Correlator dump time (sec)
\newcommand{\compt}{t_{\mathrm{comp}}} % Computation time
\newcommand{\gridkernupdate}{t_\mathrm{kernel,update}} % Update timescale for gridding kernel
\newcommand{\lsmupdate}{t_\mathrm{LSM}} % Update timescale for the LSM
\newcommand{\obst}{t_{\mathrm{obs}}} % Observation time
\newcommand{\snapshot}{t_\mathrm{snap}} % Snapshot duration
\newcommand{\minsnapshot}{t_\mathrm{snap,min}} % Minimum duration of a snapshot in which w-correction is not dominated by Earth's curvature
\newcommand{\optsnapshot}{t_\mathrm{snap,opt}} % Optimum snapshot duration (sec)
\newcommand{\subint}{t_\mathrm{subint}} % Subintegration time
\newcommand{\uv}{uv} % Visibility coordinates
\newcommand{\uvw}{u,v,w} % Visibility coordinates
\newcommand{\w}{w} % Visibility w coordinate
\newcommand{\nbytepervis}{N_\mathrm{byte,vis}} %Memory per individual visibility 
\newcommand{\maxwdev}{\Delta w_\mathrm{max}} % Maximum deviation of the w
                                % coordinate due to the Earth's curvature
\newcommand{\minwdev}{\Delta w_\mathrm{min}} % Minimum deviation of the w
                                % coordinate due to the Earth's curvature
\newcommand{\spfiltm}{\mathbb{W}_\mathrm{filt}} % Spatial filter matrix
\newcommand{\fres}{\Delta f} % Frequency resolution
\newcommand{\tressmear}{t_\mathrm{smear}} % Smearing Time resolution (sec)
\newcommand{\tres}{t_{res}} % Time resolution (sec)
\newcommand{\fracconvfuncamp}{\epsilon_\omega} % Fraction of maximal amplitude of envelope of convolution function
\newcommand{\compeff}{\eta_\mathrm{comp}} % Computational efficiency
\newcommand{\fracfband}{\eta_f} % Fraction of frequency band
\newcommand{\beamsq}{\theta_\mathrm{bs}} % Beam squint
\newcommand{\fovdiam}{\theta_\mathrm{FoV}} % Angular diameter of the field of view (rad)
\newcommand{\pixres}{\theta_\mathrm{pix}} % Angular resolution of pixel size (rad)
\newcommand{\beamdiam}{\theta_\mathrm{PSF}} % Angular diameter of the synthesised beam (rad)
\newcommand{\wavel}{\lambda} % Wavelength
\newcommand{\maxwavel}{\lambda_\mathrm{max}} % Maximum wavelength
\newcommand{\minwavel}{\lambda_\mathrm{min}} % Minimum wavelength
\newcommand{\maxwavelsubb}{\lambda_{\mathrm{sub,max}}} % Maximum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
\newcommand{\minwavelsubb}{\lambda_\mathrm{sub,min}} % Minimum wavelength of an imaging sub-band (channels over which image and pixel size are matched)
\newcommand{\earthrot}{\Omega_\mathrm{E}} % Angular velocity of the Earth (7.27 10-5 rad/sec)
\newcommand{\epsilonf}{\epsilon_f}  %Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging
% Other command definitions
\newcommand{\Qkernel}{Q_\mathrm{kernel}} %Convolution kernel quality factor, for kernel re-use in frequency
\newcommand{\bs}{\texttt{\char`\\}}

\begin{document}

% load automatic pages
\tccfrontpage

\sdptableofcontents

% Add here the abbreviations used in your document
\sdplistofabbreviations
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
    \item[CDR] Critical Design Review \vspace{-0.2cm}
    \item[CPU] Central Processing Unit\vspace{-0.2cm}
    \item[CSP] Central Signal Processor\vspace{-0.2cm}
    \item[CUFFT] \textcolor{red}{Define it.}\vspace{-0.2cm}
    \item[DFT] \textcolor{red}{Only used once, in a title. Remove/expand?}\vspace{-0.2cm}
    \item[DM] \textcolor{red}{Define it.}\vspace{-0.2cm}
    \item[EoR] Epoch of Reionisation\vspace{-0.2cm}
    \item[FFT] Fast Fourier Transform\vspace{-0.2cm}
    \item[FLOP] Floating Point Operations \vspace{-0.2cm}
    \item[FLOPS] Floating Point Operations per Second \vspace{-0.2cm}
    \item[FoV] Field of View\vspace{-0.2cm}
    \item[GPU] Graphics Processing Unit \vspace{-0.2cm}
    \item[ICD] Interface Control Document\vspace{-0.2cm}
    \item[LMA] Levenberg-Marquardt Algorithm \vspace{-0.2cm}
    \item[LOFAR] LOw Frequency ARray\vspace{-0.2cm}
    \item[LSM] Local Sky Model\vspace{-0.2cm}
    \item[MS-MFS] Multi-Scale Multi-Frequency Synthesis\vspace{-0.2cm}
    \item[NVRAM] \textcolor{red}{Define it.} \vspace{-0.2cm}
    \item[OCLD] Optimised Candidate List and Data \vspace{-0.2cm}
    \item[PDR] Preliminary Design Review\vspace{-0.2cm}
    \item[RFI] Radio Frequency Interference\vspace{-0.2cm}
    \item[PSF] \textcolor{red}{Define it.} \vspace{-0.2cm}
    \item[PSS] Pulsar Search\vspace{-0.2cm}
    \item[PST] Pulsar Timing\vspace{-0.2cm}
    \item[SDP] Scientific Data Processor\vspace{-0.2cm}
    \item[SKA] Square Kilometre Array\vspace{-0.2cm}
    \item[SKAO] SKA Organisation\vspace{-0.2cm}
    \item[WCS] \textcolor{red}{Define it.} \vspace{-0.2cm}
%    \item[] \vspace{-0.2cm}
\end{basedescript}

\noindent \emph{Reminder:}\\
Computational Cost = Number of FLOP that need to be executed to turn x into y (a.k.a.\ Computational Intensity)\\
Computational Rate = Number of FLOP executed per unit of time -- FLOPS\\
Computational Power = Computational Rate / Efficiency -- FLOPS\\
Multiply and Add = 2 Flops i.e. no Fused multiply and add assumed \\
Complex Multiply and Add = 6 Flops i.e. not o Fused multiply and add assumed\\

% Add here the symbols used in your document
\clearpage \sdplistofsymbols
\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
\item[$\viscompr$] Compression factor of visibilities for grdding for facets
  \textcolor{red}{This symbol is not used in the document, consider removing.}
  \vspace{-0.2cm}
\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
\item[$\da$] Extent of aperture \vspace{-0.2cm}
\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
\item[$\reff$] Reference frequency\vspace{-0.2cm}
\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
  coverage is almost filled\vspace{-0.2cm}
\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
  gains\vspace{-0.2cm}
\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
\item[$\nbins$] Number of bins \vspace{-0.2cm}
\item[$\ntel$] Number of antennas or stations in the array
  \vspace{-0.2cm}
\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
  axis) \vspace{-0.2cm}
\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
\item[$\navg$] Number of averages\vspace{-0.2cm}
\item[$\nbeam$] Number of beams simultaneously observed by the array
  \vspace{-0.2cm}
\item[$\nbits$] Number of bits\vspace{-0.2cm}
\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
\item[$\nbl$] Number of baselines \vspace{-0.2cm}
\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
\item[$\nbytepervis$] Number of bytes per individual visibility

\item[$\nch$] Number of channels \vspace{-0.2cm}
\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
\item[$\ngridp$] Number of grid points along one axis for the gridding
  convolution function \vspace{-0.2cm}
\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
\item[$\nfchcorr$] Number of frequency channels output by the correlator
  \vspace{-0.2cm}
\item[$\nfchgrid$] Number of frequency channels to be gridded
  \vspace{-0.2cm}
\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
  \vspace{-0.2cm}
\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
\item[$\nflop$] Number of floating point operations
  \vspace{-0.2cm}
\item[$\nflopsamp$] Number of floating point operations per sample
  \vspace{-0.2cm}
\item[$\nflopvis$] Number of floating point operations per visibility
\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
  dimension) \vspace{-0.2cm}
\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
  axis) \vspace{-0.2cm}
\item[$\nit$] Number of iterations \vspace{-0.2cm}
\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
  M\"uller terms \vspace{-0.2cm}
\item[$\nops$] Number of operations \vspace{-0.2cm}
\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
  decorrelation \vspace{-0.2cm}
\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
\item[$\nsource$] Number of sources \vspace{-0.2cm}
\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
  \vspace{-0.2cm}
\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
\item[$\qfacreuse$] Factor to allow for the reuse of functions between
  neighbouring frequency channels \vspace{-0.2cm}
\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
  field of view is than the first zero of the Airy function \vspace{-0.2cm}
\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
  multi-scale CLEAN\vspace{-0.2cm}
\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
  synthesised beam \vspace{-0.2cm}
\item[$\qfac$] Quality factor \vspace{-0.2cm}
\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
\item[$\covm$] Covariance matrix \vspace{-0.2cm}
\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
  \vspace{-0.2cm}
\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
\item[$\compt$] Computation time \vspace{-0.2cm}
\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
\item[$\obst$] Observation time \vspace{-0.2cm}
\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
  not dominated by the Earth's curvature \vspace{-0.2cm}
\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
\item[$\subint$] Sub-integration time\vspace{-0.2cm}
\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
  curvature \vspace{-0.2cm}
\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
  curvature \vspace{-0.2cm}
\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
\item[$\fres$] Frequency resolution \vspace{-0.2cm}
\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
  convolution function \vspace{-0.2cm}
\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
\item[$\beamsq$] Beam squint \vspace{-0.2cm}
\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
  \vspace{-0.2cm}
\item[$\wavel$] Wavelength \vspace{-0.2cm}
\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
  over which image and pixel size are matched) \vspace{-0.2cm}
\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
  over which image and pixel size are matched) \vspace{-0.2cm}
\item[$\earthrot$] Angular velocity of the Earth
  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
\end{basedescript}

\sdplistoffigures

\sdplistoftables

% Add here the executive summary
\sdpsummary

\newpage

% Add to the table the list of applicable and reference documents (this is NOT
% meant for your usual bibiliography, only for SKA docs)
\sdpreferencedocs

\subsection*{Applicable Documents}

The following documents are applicable to the extent stated herein. In the
event of conflict between the contents of the applicable documents and this
document, \emph{the applicable documents} shall take precedence.

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
     {\bf [AD01]} & Dewdney, P. E. (2013). SKA1 System Baseline Design. SKA Office\\
     {\bf [AD02]} & McCool, R., Cornwell, T. (2013). Miscellaneous Corrections
     to the Baseline Design\\
     {\bf [AD03]} & SKA Phase 1 System (Level 1) Requirements Specification, T.Cornwell, SKA-OFF.SE.ARC-SKO-SRS-001\_2\\
     {\bf [AD04]} & PDR.01 SDP.ARCH document\\
     {\bf [AD05]} & SDP Costing spreadsheet\\
     {\bf [AD06]} & Cornwell, T.J. (2015). SKA1 Telescope Calibration
     Framework. SKA Office (draft version)\\
     {\bf [AD07]} & CSP--SDP ICD\\
     \hline
   \end{tabularx}}
\end{center}

\subsection*{Reference Documents}

The following documents are referenced in this document. In the event of
conflict between the contents of the referenced documents and this document,
\emph{this document} shall take precedence.

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
     {\bf [RD01]} & SKA-TEL-SDP-0000041, F. Malan: iPython Performance Model Description\\
    {\bf [RD02]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
    {\bf [RD03]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
    {\bf [RD04]} & SKA-TEL-SDP-0000028, A.-J. Boonstra: Ingest Pipeline\\
    {\bf [RD05]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
    {\bf [RD06]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
    {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
    {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent
    Averaging\\
    {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
    {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
    {\bf [RD11]} & Alexander et al., Software and Computing CoDR, Analysis of requirements derived from the DRM, WP2-050.020.010-RR-001\\
    {\bf [RD12]} & SKA-TEL-SDP-0000016, R. Bolton: SDP Archive Size
    Estimates\\
    {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
    {\bf [RD14]} & \url{https://confluence.ska-sdp.org/pages/viewpage.action?title=MS+in+UML&spaceKey=ARCH}\\
    {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821,
    \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
    {\bf [RD16]} & \url{http://www.lofar.org/operations/lib/exe/fetch.php?media=:public:documents:casa_image_for_lofar_0.03.00.pdf}\\
     \hline
   \end{tabularx}}
\end{center}

 \begin{center}{
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
    {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis
    imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
    {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
    {\bf [RD19]} & Cornwell, T. J., Voronkov, M. A., Humphreys, B. (2012). Wide field imaging for the Square Kilometre Array.\\
    {\bf [RD20]} & An Efficient Work-Distribution Strategy for Gridding Radio-Telescope Data on GPUs, John W. Romein, ICS’12, June 25–29, 2012, San Servolo Island, Venice, Italy. Source code available at: \url{ftp://ftp.astron.nl/outgoing/romein/Gridding-0.2.tar.bz}\\
   {\bf [RD21]} & Mitchell, D. a. (2014). Analysis of w-projection kernel size. SKA SDP Consortium\\
   {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
   {\bf [RD23]} & A parametric model for the calibration and imaging costs of SKA1, T.J. Cornwell, 12 Feb 2013, SKA Draft Memo\\
     {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
   {\bf [RD25]} & Tasse, C. a. (2013). Applying full polarization A-Projection to very wide field of view instruments: An imager for LOFAR. A\&A.\\
   {\bf [RD26]} & Faceted imaging in AIPS, Kogan and Greisen, AIPS Memo 113, 2009\\
   {\bf [RD27]} & Radio-interferometric imaging of very large fields - The
   problem of non-coplanar arrays, Cornwell, T. J.; Perley, R. A., Astronomy
   and Astrophysics (ISSN 0004-6361), vol. 261, no. 1, p. 353-364\\
    {\bf [RD28]} & Wide Field Imaging at Low Frequencies, Sault, R. J.,  The Universe at Low Radio Frequencies, Proceedings of IAU Symposium 199, held 30 Nov - 4 Dec 1999, Pune, India. Edited by A. Pramesh Rao, G. Swarup, and Gopal-Krishna, 2002., p.508\\
    {\bf [RD29]} & Roofline: An Insightful Visual Performance Model for
    Multicore Architectures, Williams, Waterman and Patterson,
    DOI:10.1145/1498765.1498785, April 2009, Vol. 52, No. 4, Communications of
    the ACM\\
    {\bf [RD30]} & Whiting, M., 2012, Duchamp: a 3D source finder for spectral-line data, Mon. Not. R. Astron. Soc., Volume 421, Issue 4,  pages 3242–3256, April 2012 \\
    {\bf [RD31]} & Hales, C. A., Murphy, T., Curran, J. R., Middelberg, E., Gaensler, B. M., Norris, R. P., blobcat: software to catalogue flood-filled blobs in radio images of total intensity and linear polarization, Mon. Not. R. Astron. Soc., Volume 425,  Issue 2, pages 1365-2966, September 2012\\
    {\bf [RD32]} & Hancock, P.J., Murphy, T., Gaensler, B.M., Hopkins, A.,
    Curran, J.R., (2012), Compact continnum source finding for next generation
    radio surveys, Mon. Not. R. Astron. Soc. 422, 1812–1824\\
   {\bf [RD33]} & Whiting, M.,  Humphreys, B., (2012) Source-Finding for the
   Australian Square Kilometre Array Pathfinder, Publications of the
   Astronomical Society of Australia, Volume 29, Special Issue 03, 2012,
   pp 371-381.\\
   {\bf [RD34]} Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD35]} Conway, J. E., Cornwell, T. J., & Wilkinson, P. N. (1990). Multi-Frequency Synthesis - a New Technique in Radio Interferometric Imaging. Monthly Notices of the Royal Astronomical Society, 246, 490.\\

   \hline
 \end{tabularx}}
\end{center}

% The actual content goes here
\newpage
\section{Introduction}

\subsection{Purpose of the document}

The purpose of this document is to describe the Multi-Scale MultiFrequency Synthesis algorithm, the variant implementations in CASA and ASKAPsoft, and the scaling for insertion in the SDP Performance Model.

\subsection{Scope of the document}

\clearpage
\section{Overview of MSMFS}
\label{sec:overview}

MS-MFS was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g. [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data. The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law.

\clearpage
\section{Major cycle}
\label{sec:major}
\subsection{Computational task}

\subsection{ASKAP implementation}

\subsection{CASA implementation}

\clearpage
\section{Minor cycle}
\label{sec:minor}

\subsection{Computational task}

\subsection{ASKAP implementation}

\subsection{CASA implementation}

\clearpage
\section{A projection}
\label{sec:minor}

\clearpage
section{Wide-band behaviour}
\label{sec:wideband}

\begin{itemize}
\item Does naive broadband work?
\item Bhatnagar et al WB algorithm	
\end{itemize}

\clearpage
\section{Implementation in SDP}
\label{sec:implementation}

\clearpage
\section{Resource usage}
\label{sec:resource}

% \label{section:main}

% Add the bibliography
\clearpage \addcontentsline{toc}{section}{References}
\bibliography{your_bibfile}%

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: